Landau quantization of nearly degenerate bands

### 2018-09-19 paper.tex
progress

$(l^2\delta S)^{-1}$ is a measure of the field-induced hybridization of the quasidegenerate orbits, and shall be called the hybridization parameter. 
In the regime that $l^2\delta S$ is much greater than the Zeeman interaction \red{[Zeeman interaction? Why not just 1.]}, we expect that the Onsager-Lifshitz quantization rules apply independently to the two concentric (but decoupled) orbits. In the opposite regime, we should recover the known quantization rule for degenerate orbits. [topoferm, 100p, roth, mikitik] Our goal is to present a unifying rule that interpolates smoothly between the two known regimes.

@@

Let us apply a magnetic field, and first consider the spin-degenerate Landau levels obtained by neglecting both $\delta H$ and the Zeeman splitting. In the  semiclassical description,  we are interested in  field-induced dynamics within a low-energy degenerate band of $\hat{H}_0$ with  dispersion $\var(\bk)$, e.g., $\var(\bk){=}\hbar^2 k^2/2m{+}\order(k^4)$ \red{[What is contained in this $\order(k^4)$?]}in the Rashba model.    Landau quantization is well-known to derive from the Peierls substitution $\var(\bk){\rightarrow}\var(\bK)$, [Peierls] with the kinematic quasimomentum operator   $\boldsymbol{K}{:}{=}\boldsymbol{k}{+}(e/\hbar) \boldsymbol{A}(i\nabla_{\boldsymbol{k}})$ and $\boldsymbol{A}(\br)$ being the electromagnetic vector potential. Assuming that the field is aligned in the ${-}z$ direction, $[K_x,K_y]{=}i\lmt$. The semiclassical solution corresponds to  cyclotron revolutions of an electron around a constant-energy contour of $\var(\bk)$ [see Fig. \ref{fig:orbits}(b)]]; this contour, equipped with an orientation given by the equation of motion ($ \hbar dk_{\alpha}/dt {=} \lmt \epsilon_{\ab}\partial \var/\partial k_{\beta} $, with $\epsilon_{xy}{=}{-}\epsilon_{yx}{=}1$), shall be referred as the zeroth-order orbit $\frako_0$.  In this work, we focus only on closed orbits which do not wrap 
non-contractibly around the Brillouin torus.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{(a) Trefoille-shaped cyclotron orbit for the Rashba-Zeeman model at the energy of the Dirac point. The orbit is drawn over the  {spin-orbit field} ($d_x,d_y$) of the Hamiltonian $H_{RZ}{=}d_0(\bk)I{+}d_x(\bk)\sigma_x{+}d_y(\bk)\sigma_y$, with $d_j$ obtainable from \q{eq:RZ-Hamiltonian}. The wavefunction along the inner (resp.\ outer) component of the trefoille orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-orbit field. (b) Black dashed lines illustrate a zeroth-order approximation of the Landau levels (assuming Klein tunneling); the leading-order correction is indicated by red lines. The Dirac point lies at $10$meV. Parameters for (b): $m$: 0.02$\hbar^2$/eV$\cdot$\AA$^2$, $\alpha$: 0.05eV$\cdot$\AA, $Z_\parallel$: 0.001eV. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. (d) For various configurations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in left panel $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}), where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. Corresponding $l^2 \delta S^2/S$ are plotted in right panel. We only plotted in region where $0.4<\epsilon_0<1.0$(eV).\footnote{$l^2S\gg 1$ is assured by requiring $\epsilon_0>0.4$; the reason of choosing $\epsilon_0<1.0$ is Dirac point lying at high energy requires too much harmonics in numerical diagonalization. This requirement is equivalent to $\sqrt{0.8m}\alpha/\hbar<Z<\sqrt{2m}\alpha/\hbar$, corresponding to the boundaries of the plots in (d).} Parameters for (d): $m$: 0.01$\hbar^2$/eV$\cdot$\AA$^2$, $l=100$\AA. \textbf{please choose formatting such that footnote is at the back} \label{fig:RZ}}
\end{figure}

@@

As alluded to in \s{sec:qtznrules}, the adiabatic limit of two independent orbits does not exist if the two spin-orbit-split orbits exactly intersect at a II-Dirac point. To investigate this failure of adiabaticity, let us study the Landau levels at energies close to the II-Dirac point ($\epsilon_0$). We focus first on the regime where, on average (over $\frako_0$), the energy splitting $|\delta \var_+{-}\delta \var_-|$ (due to both  spin-orbit and in-plane Zeeman interactions) is much greater than the interband, out-of-plane Zeeman interaction $B_{\perp}|\calm_{\pm}|$. This implies that the split orbits do not appreciably hybridize -- except in the vicinity of the II-Dirac point where orbits (nearly) touch. Near this II-Dirac point, the dynamics of a two-component wavepacket is described 
 by the effective Hamiltonian
\e{
\calh = \alpha k_{E} \,\omega^{\sma{\perp}}_c t\,\sigma_x+(Z_\parallel-\alpha k_{E})\sigma_y-Z_\perp\sigma_z,\la{effhamIIDirac}
}
with $k_E{=}\sqrt{2mE}/\hbar$ and  $t$ a time variable which parametrizes the zeroth-order orbit:
$\bk(t){=}({-}k_E \cos \omega_c^{\sma{\perp}} t, k_E \sin \omega_c^{\sma{\perp}} t ).$ 
$\calh$ above has been linearized with respect to  $t$, which vanishes in the vicinity of the Dirac point. The eigenbasis of $\sigma_x$ corresponds to two diabatic levels with characteristic slope $v_d{:}{=}\alpha k_E\omega^{\sma{\perp}}_c$; near $t{=}0$, the diabatic levels anticross  due to the hybridization terms proportional to $\sigma_{y,z}$. Generally, the probability of tunneling across the hybridization-induced barrier is  $\rho^2{=}\exp(-2\pi\barmu)$, with $\barmu{=}{E_g}^2/{2v_d \hbar}$ and $E_g$ the energy gap at the center of the anticrossing;[Landau and lifshitz, Wittig] in this context,
\e{\barmu=\frac{Z_\perp^2+(Z_\parallel-\alpha k_{E})^2}{2\alpha k_{E}\var^{\sma{\perp}}_c}, \la{mu1}}
with $2\alpha k_E$ the spin-orbit splitting at energy $E$ [cf.\ \fig{fig:orbits}(a)]. In the absence of the out-of-plane Zeeman coupling ($Z_\perp{=}0$), $\bar{\mu}$ would vanish at the energy of the II-Dirac point -- leading to unit-probability Klein tunneling, as was first proposed by [Obrien] in a different model realization of the II-Dirac point. However, with a proper account of $Z_\perp$, we find instead that $\bar{\mu}$ has a nonvanishing minimum: $\bar{\mu}_{\sma{\text{min}}}{:}{=} (m/m_0)(Z_\perp/2\alpha k_{E})$, which is a product of the dimensionless effective mass and the ratio of the out-of-plane Zeeman splitting over the spin-orbit splitting. This minimum is practically negligible for  semiconductor heterostructures with $m{\ll}m_0$, but may be relevant in heavy-fermion systems.

To determine the Landau levels, one needs not just the probability for tunneling but also the phase of the probability amplitude. The full information is encoded in a scattering matrix [AALG]
\e{\mathbb{S}=\matrixtwo{\tau e^{i\barphi}}{-\rho}{\rho}{\tau e^{-i\barphi}},\;\rho=e^{-\pi\barmu},\;\tau=\sqrt{1-\rho^{2}}\la{scattmat}}
which is a connection formula that matches the two-component semiclassical (WKB) wavefunction across the Dirac point. $\mathbb{S}$ in \q{scattmat} has been expressed in the basis of spin-split energy bands, so  the diagonal elements of $\mathbb{S}$ are amplitudes for intraband (adiabatic) transmission; the phase of this amplitude is $\barphi{:}{=}\barmu{-}\barmu\ln\barmu{+}\text{arg}[\Gamma(i\barmu)]{+}\pi/4$ with $\Gamma$ the gamma function.

Away from the II-Dirac point, a semiclassical wavepacket undergoes adiabatic (band-conserving) dynamics. The time-ordered, coherent process of interband tunneling and adiabatic dynamics is described by the propagator:
\e{{\A}= \matrixtwo{\tau e^{i\bar{\varphi}}}{-\rho}{\rho}{\tau e^{-i\bar{\varphi}}} \diagmatrix{e^{i\tilde{\lambda}_+}}{e^{i\tilde{\lambda}_-}}. \la{propinplanezeeman}}
$\tilde{\lambda}_{\pm}$ is generated by $\H_{\pm \pm}$ and has the same form as \q{phaseindependentorbit}, with $B\int_0^{T_c}\calm_{\pm\pm}dt/\hbar$ equal (in this model) to the single-band Berry phase ($\phi_{\pm}^B$) of band $\pm$ on $\frako_0$. At energies far above the Dirac point ($|E{-}\epsilon_0|{\gg}|Z_{\perp}|$) where the Zeeman interaction is negligible, both orbits encircle the Dirac point and hence $\phi_{\pm}^B{\approx}{\pm}\pi$; far below the Dirac point, neither orbit encircles the Dirac point, leading to $\phi_{\pm}^B{\approx}0$. At intermediate energies, $\phi_{+}^B$ (resp.\ $\phi_{-}^B$) is a continuous and increasing (resp.\ decreasing) function of energy, as supported by the Bloch-sphere argument in Fig. \ref{fig:blochsphere}.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{blochsphere.png}
\caption{Energy dependence of the Berry phase for inner (upper panel) and outer (lower panel) orbits. For two band models, the Berry phase equals half the solid angle enclosed by the loop swept by spin orbit field on the unit sphere. The solid angles are indicated by blue outlines and the energy is measured from $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}).\label{fig:blochsphere}}
\end{figure}

Our quantization rule [\qq{eq:rule}{eq:H1}], with $\A$ having the form of \q{propinplanezeeman}, may be equivalently expressed as
\e{
\text{cos}\left[\frac{\Omega_{-}+\Omega_{+}}{2}\right]=\tau\,\text{cos}\left[\frac{\Omega_{-}-\Omega_{+}}{2}+\bar{\varphi}\right], \lin
\ins{with}\;\Omega_{\pm}=l^2 (S\pm \delta S/2)+\phi_{\pm}^B +\gamma_{\pm}, \la{coscos}
}
$\Omega_{\pm}$ is the phase acquired by an electron in adiabatically traversing orbit $\pm$; the quantum Maslov correction $\gamma_{\pm}{=}\pi$ for circular orbits. A quantization rule analogous to \q{coscos} was previously derived by us for a II-Dirac point that intermediates an electron- and hole-like orbit [AALG].

@@

In the limit $B{\rightarrow}0$ (equivalently, $\bar{\mu}{\rightarrow} \infty$), Landau-Zener tunneling is negligible ($\tau{\rightarrow}1,\bar{\varphi}{\rightarrow}0$), and \q{coscos} simplifies to the Onsager-Lifshitz-Roth rules ($\Omega_{\pm}/2\pi{\in}\Z$) for  two independent orbits. This simplification occurs even at the energy of the Dirac point, where  the intraband amplitude $\tau$ has a nonvanishing minimum with respect to energy: $\tau_{\sma{\text{min}}}{\approx}\sqrt{2\pi\bar{\mu}_{\sma{\text{min}}}}$, due to the out-of-plane Zeeman coupling.

For any nonzero field, $0{<}\tau{<}1$ implies that the two incommensurate harmonics $(\Omega_+{\pm}\Omega_-)$ in \q{coscos} compete to produce a quasirandom Landau-level spectrum [kaganovslutskin]. That is to say, the spectrum is disordered on the scale of nearest-neighbor level spacings but has longer-ranged correlations. To appreciate this, let us analyze the Landau levels (near the Dirac point) perturbatively in the small parameter $\tau$. To zeroth order, \q{coscos} reduces to\footnote{Notice that $\phi_+ + \phi_-=0$ since for any two band model, Berry phases of the two bands cancel on the same orbit ($\frako_0$ here).} 
\e{2l^2S(E_n^0)+\pi = 2\pi n \rightarrow \{E_n^0(B)\}_{n\in\Z}.\label{trefoillerule}} 
 This may be interpreted as a quantization rule for a trefoille-shaped orbit with net action $2S{=}4\pi m E{/\hbar^2}$; the corresponding Landau levels, as plotted with dashed lines in Fig.\ \ref{fig:RZ}(b), are  periodic in $E{\rightarrow}E{+}\pi/l^2(\partial S/\partial E)$ and in $l^2{\rightarrow}l^2{+}\pi/S$. Both periods are approximately half that of either disconnected spin-orbit-split orbit, in the assumed quasidegenerate regime $\delta S{\ll}S$. The $\pi$ correction in \q{trefoillerule} is the Berry phase associated to a $2\pi$ rotation of the wavefunction spin over a single trefoille orbit, as illustrated in Fig. \ref{fig:RZ}(a). The Maslov phase, which is generally equal to $\pi$ times the rotation number of an orbit, is trivially $2\pi$ for the trefoille orbit.

@@

\textbf{[Rest of section needs work]
} In the regime of quasidegenerate orbits ($\Delta/E{\ll}1$) with intermediate adiabaticity ($\Delta/\epsilon_c{\sim}1$),  our quantization rule \qq{eq:rule}{eq:H1} remains valid but the propagator $\cala$ no longer has the closed analytic form of \q{propinplanezeeman}. We may anyway simulate $\cala$ numerically, and compare the semiclassically-obtained Landau levels with the exact solution of $H_{RZ}(\boldsymbol{K})$ (see Fig. \ref{fig:RZ} (c-d)). This exact solution is obtained by large-scale numerical diagonalization [appendix], since (to our knowledge) there exists no analytic solution for generic values of $\alpha,\zpar$ and $B$. \textbf{[Say a few words about the comparison]}

For the special case of $\alpha=0$, the spectra of $H_{RZ}(\boldsymbol{K})$ and  $\mathcal{A}[\mathfrak{o}]$ are both exactly soluble, leading to identical Landau levels with phase offset $\lambda_{\pm}=\pm 2\pi ml^2\sqrt{Z^2+g_s'^2/16m^2l^4}$.\textbf{[Let's be more specific about the form of the spectrum.]}

@@

To understand why rotational symmetry reduces codimension in the Landau-level problem, we return to the Rashba model [cf. \ \q{eq:Rashba-Hamiltonian}] absent Dresselhaus spin-orbit coupling and in-plane Zeeman field. This model has a continuous rotational symmetry (denoted $\mathfrak{c}_{\infty}$), which manifests as the constancy of $\calh(\bk){:}{=}\sum_{j=1}^3r_j\tau_j$ over the circular zeroth-order orbit [cf.\ \q{rashbaeffham}]; here, $\br{=}(r_1,r_2,r_3)$ is a real three-vector and $\{\tau_j\}$ a set of Pauli matrices.  Consequently, the propagator  simplifies to  the exponential form:
\e{\A{=}e^{-i2\pi\sum_{j}r_j\tau_j}, \;\;(r_1,r_3){=}\bigg(\f{g_{\sma{\ext{eff}}}}{4}{-}\f1{2},\f1{2}{-}\f{\alpha k_{\sma{E}}}{\var_c}\bigg).\la{expform}} 
$r_2{=}0$ in the Rashba model, but nonzero $r_2$ corresponds to a symmetry-allowed deformation that we allow for greater generality -- any $SU(2)$ matrix can be expressed in the exponential form [cf.\ \q{expform}]  for some (possibly non-unique) $\br$, which are the symmetric  coordinates.

>>>>

$(l^2\delta S)^{-1}$ is a measure of the field-induced hybridization of the quasidegenerate orbits, and shall be called the hybridization parameter. 
In the regime that $l^2\delta S$ is much greater than the Zeeman interaction, we expect that the Onsager-Lifshitz quantization rules apply independently to the two concentric (but decoupled) orbits. In the opposite regime, we should recover the known quantization rule for degenerate orbits. [topoferm, 100p, roth, mikitik] Our goal is to present a unifying rule that interpolates smoothly between the two known regimes.

@@

Let us apply a magnetic field, and first consider the spin-degenerate Landau levels obtained by neglecting both $\delta H$ and the Zeeman splitting. In the  semiclassical description,  we are interested in  field-induced dynamics within a low-energy degenerate band of $\hat{H}_0$ with  dispersion $\var(\bk)$, e.g., $\var(\bk){=}\hbar^2 k^2/2m{+}\order(k^4)$ in the Rashba model.    Landau quantization is well-known to derive from the Peierls substitution $\var(\bk){\rightarrow}\var(\bK)$, [Peierls] with the kinematic quasimomentum operator   $\boldsymbol{K}{:}{=}\boldsymbol{k}{+}(e/\hbar) \boldsymbol{A}(i\nabla_{\boldsymbol{k}})$ and $\boldsymbol{A}(\br)$ being the electromagnetic vector potential. Assuming that the field is aligned in the ${-}z$ direction, $[K_x,K_y]{=}i\lmt$. The semiclassical solution corresponds to  cyclotron revolutions of an electron around a constant-energy contour of $\var(\bk)$ [see Fig. \ref{fig:orbits}(b)]]; this contour, equipped with an orientation given by the equation of motion ($ \hbar dk_{\alpha}/dt {=} \lmt \epsilon_{\ab}\partial \var/\partial k_{\beta} $, with $\epsilon_{xy}{=}{-}\epsilon_{yx}{=}1$), shall be referred as the zeroth-order orbit $\frako_0$.  In this work, we focus only on closed orbits which do not wrap 
non-contractibly around the Brillouin torus.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{(a) Trefoille-shaped cyclotron orbit for the Rashba-Zeeman model at the energy of the Dirac point. The orbit is drawn over the  {spin-orbit field} ($d_x,d_y$) of the Hamiltonian $H_{RZ}{=}d_0(\bk)I{+}d_x(\bk)\sigma_x{+}d_y(\bk)\sigma_y$, with $d_j$ obtainable from \q{eq:RZ-Hamiltonian}. The wavefunction along the inner (resp.\ outer) component of the trefoille orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-orbit field. (b) Black dashed lines illustrate a zeroth-order approximation of the Landau levels (assuming Klein tunneling); the leading-order correction is indicated by red lines. The Dirac point lies at $10$meV. Parameters for (b): $m$: 0.02$\hbar^2$/eV$\cdot$\AA$^2$, $\alpha$: 0.05eV$\cdot$\AA, $Z_\parallel$: 0.001eV. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. (d) For various configurations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in left panel $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}), where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. Corresponding $l^2 \delta S^2/S$ are plotted in right panel. We only plotted in region where $0.4<\epsilon_0<1.0$(eV).\footnote{$l^2S\gg 1$ is assured by requiring $\epsilon_0>0.4$; the reason of choosing $\epsilon_0<1.0$ is Dirac point lying at high energy requires too much harmonics in numerical diagonalization. This requirement is equivalent to $\sqrt{0.8m}\alpha/\hbar<Z<\sqrt{2m}\alpha/\hbar$, corresponding to the boundaries of the plots in (d).} Parameters for (d): $m$: 0.01$\hbar^2$/eV$\cdot$\AA$^2$, $l=100$\AA. \textbf{[please choose formatting such that footnote is at the back]}\red{[Sorry, I haven't figured out how to do this. I will do this later.]} \label{fig:RZ}}
\end{figure}

@@

As alluded to in \s{sec:qtznrules}, the adiabatic limit of two independent orbits does not exist if the two spin-orbit-split orbits exactly intersect at a II-Dirac point. To investigate this failure of adiabaticity, let us study the Landau levels at energies close to the II-Dirac point ($\epsilon_0$). We focus first on the regime where, on average (over $\frako_0$), the energy splitting $|\delta \var_+{-}\delta \var_-|$ (due to both  spin-orbit and in-plane Zeeman interactions) is much greater than the interband, out-of-plane Zeeman interaction $B_{\perp}|\calm_{\pm}|$. This implies that the split orbits do not appreciably hybridize -- except in the vicinity of the II-Dirac point where orbits (nearly) touch. Near this II-Dirac point, the dynamics of a two-component wavepacket is described 
 by the effective Hamiltonian
\e{
\calh = \alpha k_{E} \,\omega^{\sma{\perp}}_c t\,\sigma_x+(Z_\parallel-\alpha k_{E})\sigma_y-Z_\perp\sigma_z,\la{effhamIIDirac}
}
with $k_E{=}\sqrt{2mE}/\hbar$ and  $t$ a time variable which parametrizes the zeroth-order orbit:
$\bk(t){=}({-}k_E \cos \omega_c^{\sma{\perp}} t, k_E \sin \omega_c^{\sma{\perp}} t ).$ 
$\calh$ above has been linearized with respect to  $t$, which vanishes in the vicinity of the Dirac point. The eigenbasis of $\sigma_x$ corresponds to two diabatic levels with characteristic slope $v_d{:}{=}\alpha k_E\omega^{\sma{\perp}}_c$; near $t{=}0$, the diabatic levels anticross  due to the hybridization terms proportional to $\sigma_{y,z}$. Generally, the probability of tunneling across the hybridization-induced barrier is  $\rho^2{=}\exp(-2\pi\barmu)$, with $\barmu{=}{E_g}^2/{2v_d \hbar}$ and $E_g$ the energy gap at the center of the anticrossing;[Landau and lifshitz, Wittig] in this context,
\e{\barmu=\frac{Z_\perp^2+(Z_\parallel-\alpha k_{E})^2}{2\alpha k_{E}\var^{\sma{\perp}}_c}, \la{mu1}}
with $2\alpha k_E$ the spin-orbit splitting at energy $E$ [cf.\ \fig{fig:orbits}(a)]. In the absence of the out-of-plane Zeeman coupling ($Z_\perp{=}0$), $\bar{\mu}$ would vanish at the energy of the II-Dirac point -- leading to unit-probability Klein tunneling, as was first proposed by [Obrien] in a different model realization of the II-Dirac point. However, with a proper account of $Z_\perp$, we find instead that $\bar{\mu}$ has a nonvanishing minimum: $\bar{\mu}_{\sma{\text{min}}}{:}{=} ( m/m_0)(Z_\perp/4\alpha k_{E})$, which is a product of the dimensionless effective mass and the ratio of the out-of-plane Zeeman splitting over the spin-orbit splitting. This minimum is practically negligible for  semiconductor heterostructures with $m{\ll}m_0$, but may be relevant in heavy-fermion systems.

To determine the Landau levels, one needs not just the probability for tunneling but also the phase of the probability amplitude. The full information is encoded in a scattering matrix [AALG]
\e{\mathbb{S}=\matrixtwo{\tau e^{i\barphi}}{-\rho}{\rho}{\tau e^{-i\barphi}},\;\rho=e^{-\pi\barmu},\;\tau=\sqrt{1-\rho^{2}}\la{scattmat}}
which is a connection formula that matches the two-component semiclassical (WKB) wavefunction across the Dirac point. $\mathbb{S}$ in \q{scattmat} has been expressed in the basis of Zeeman coupling corrected energy bands, i.e., eigenvector of Eq. \ref{effhamIIDirac}, so the diagonal elements of $\mathbb{S}$ are amplitudes for intraband (adiabatic) transmission; the phase of this amplitude is $\barphi{:}{=}\barmu{-}\barmu\ln\barmu{+}\text{arg}[\Gamma(i\barmu)]{+}\pi/4$ with $\Gamma$ the gamma function.

Away from the II-Dirac point, a semiclassical wavepacket undergoes adiabatic (band-conserving) dynamics. The time-ordered, coherent process of interband tunneling and adiabatic dynamics is described by the propagator:
\e{{\A}= \matrixtwo{\tau e^{i\bar{\varphi}}}{-\rho}{\rho}{\tau e^{-i\bar{\varphi}}} \diagmatrix{e^{i\tilde{\lambda}_+}}{e^{i\tilde{\lambda}_-}}. \la{propinplanezeeman}}
In the basis of Zeeman coupling corrected energy bands, the band-conserving phases $\tilde{\lambda}_{\pm}$ consist of two parts: dynamical phase $l^2 \delta S$ with $\delta S$ now accounting for both spin-split and Zeeman-split area corrections, and single band Berry phases $\phi^B_\pm$ evaluated on $\frako_0$. At energies far above the Dirac point ($|E{-}\epsilon_0|{\gg}|Z_{\perp}|$) where the Zeeman interaction is negligible, both orbits encircle the Dirac point and hence $\phi_{\pm}^B{\approx}{\pm}\pi$; far below the Dirac point, neither orbit encircles the Dirac point, leading to $\phi_{\pm}^B{\approx}0$. At intermediate energies, $\phi_{+}^B$ (resp.\ $\phi_{-}^B$) is a continuous and increasing (resp.\ decreasing) function of energy, as supported by the Bloch-sphere argument in Fig. \ref{fig:blochsphere}.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{blochsphere.png}
\caption{Energy dependence of the Berry phases for inner (upper panel) and outer (lower panel) orbits. For two band models, the Berry phase equals half the solid angle enclosed by the loop swept by spin-orbit field on the unit sphere. The solid angles are indicated by blue outlines and the energy is measured from $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}).\label{fig:blochsphere}}
\end{figure}

Our quantization rule [\qq{eq:rule}{eq:H1}], with $\A$ having the form of \q{propinplanezeeman}, may be equivalently expressed as
\e{
\text{cos}\left[\frac{\Omega_{-}+\Omega_{+}}{2}\right]=\tau\,\text{cos}\left[\frac{\Omega_{-}-\Omega_{+}}{2}+\bar{\varphi}\right], \lin
\ins{with}\;\Omega_{\pm}=l^2 (S\pm \delta S/2)+\phi_{\pm}^B +\gamma_{\pm}, \la{coscos}
}
$\Omega_{\pm}$ is the phase acquired by an electron in adiabatically traversing orbit $\pm$; the quantum Maslov correction $\gamma_{\pm}{=}\pi$ for circular orbits. Since $\phi^B_\pm$ are evaluated both on $\frako_0$, $\phi^B_-+\phi^B_+=0$ in two band models. A quantization rule analogous to \q{coscos} was previously derived by us for a II-Dirac point that intermediates an electron- and hole-like orbit [AALG].

@@

In the limit $B{\rightarrow}0$ (equivalently, $\bar{\mu}{\rightarrow} \infty$), Landau-Zener tunneling is negligible ($\tau{\rightarrow}1,\bar{\varphi}{\rightarrow}0$), and \q{coscos} simplifies to the Onsager-Lifshitz-Roth rules ($\Omega_{\pm}/2\pi{\in}\Z$) for  two independent orbits.

For any nonzero field, $0{<}\tau{<}1$ implies that the two incommensurate harmonics $(\Omega_+{\pm}\Omega_-)$ in \q{coscos} compete to produce a quasirandom Landau-level spectrum [kaganovslutskin]. That is to say, the spectrum is disordered on the scale of nearest-neighbor level spacings but has longer-ranged correlations. To appreciate this, let us analyze the Landau levels (near the Dirac point) perturbatively in the small parameter $\tau$. To zeroth order, \q{coscos} reduces to
\e{2l^2S(E_n^0)+\pi = 2\pi n \rightarrow \{E_n^0(B)\}_{n\in\Z}.\label{trefoillerule}} 
 This may be interpreted as a quantization rule for a trefoille-shaped orbit with net action $2S{=}4\pi m E{/\hbar^2}$; the corresponding Landau levels, as plotted with dashed lines in Fig.\ \ref{fig:RZ}(b), are  periodic in $E{\rightarrow}E{+}\pi/l^2(\partial S/\partial E)$ and in $l^2{\rightarrow}l^2{+}\pi/S$. Both periods are approximately half that of either disconnected spin-orbit-split orbit, in the assumed quasidegenerate regime $\delta S{\ll}S$. The $\pi$ correction in \q{trefoillerule} is the Berry phase associated to a $2\pi$ rotation of the wavefunction spin over a single trefoille orbit, as illustrated in Fig. \ref{fig:RZ}(a). The Maslov phase, which is generally equal to $\pi$ times the rotation number of an orbit, is trivially $2\pi$ for the trefoille orbit.

@@

In the regime of quasidegenerate orbits ($\delta S/S \ll 1$) with intermediate adiabaticity ($l^2\delta S\sim 1$),  our quantization rule \qq{eq:rule}{eq:H1} remains valid but the propagator $\cala$ no longer has the closed analytic form of \q{propinplanezeeman}. We may anyway simulate $\cala$ numerically, and compare the semiclassically-obtained Landau levels with the exact solution of $H_{RZ}(\boldsymbol{K})$ (see Fig. \ref{fig:RZ} (d)). This exact solution is obtained by large-scale numerical diagonalization [appendix], since (to our knowledge) there exists no analytic solution for generic values of $\alpha,\zpar$ and $B$. Errors of \qq{eq:rule}{eq:H1} in predicting Landau levels are less than 0.3\% of cyclotron energy in Fig. \ref{fig:RZ} (d). Similar to Rashba-Dresselhaus model, the observed errors are strongly positively related to $l^2\delta S^2/S$.

For the special case of $\alpha=0$, the spectra of $H_{RZ}(\boldsymbol{K})$ and $\mathcal{A}[\mathfrak{o}]$ are both exactly soluble. The eigenphases of $\A$ are $\lambda_{\pm}=\pm 2\pi ml^2\sqrt{Z_\parallel^2/\hbar^4+g_\text{eff}^2/16m^2l^4}$. Together with Eq. (\ref{eq:rule}), $\lambda_\pm$ here generate exact the same Landau level spectrum as $H_{RZ}(\boldsymbol{K})$.

@@

To understand why rotational symmetry reduces codimension in the Landau-level problem, we return to the Rashba model [cf. \ \q{eq:Rashba-Hamiltonian}] absent Dresselhaus spin-orbit coupling and in-plane Zeeman field. This model has a continuous rotational symmetry (denoted $\mathfrak{c}_{\infty}$), which manifests as the constancy of $\calh(\bk){:}{=}\sum_{j=1}^3r_j\tau_j$ over the circular zeroth-order orbit [cf.\ \q{rashbaeffham}]; here, $\br{=}(r_1,r_2,r_3)$ is a real three-vector and $\{\tau_j\}$ a set of Pauli matrices.  Consequently, the propagator  simplifies to  the exponential form:
\e{\A{=}e^{-i2\pi\sum_{j}r_j\tau_j}, \;\;(r_1,r_3){=}\bigg(\f{g_{\sma{\text{eff}}}}{4}{-}\f1{2},\f1{2}{-}\f{\alpha k_{\sma{E}}}{\var_c}\bigg).\la{expform}} 
$r_2{=}0$ in the Rashba model, but nonzero $r_2$ corresponds to a symmetry-allowed deformation that we allow for greater generality -- any $SU(2)$ matrix can be expressed in the exponential form [cf.\ \q{expform}]  for some (possibly non-unique) $\br$, which are the symmetric  coordinates.
